# Reproducible Research: Peer Assessment 1

```{r setoptions, echo=TRUE}
opts_chunk$set(echo = FALSE)
```

## Loading and preprocessing the data
```{r read_data}
dt = read.csv("activity.csv", as.is = TRUE) 
str(dt)
summary(dt)
```

## What is mean total number of steps taken per day?
```{r num_steps_per_per_day_hist, fig.width= 11}
mean_per_day = 0
median_per_day = 0
num_steps_per_per_day_hist <- function(dt){
       steps_per_day = sapply(split(dt$steps,dt$date), sum, na.rm = TRUE)
       mean_per_day <<- round(mean(steps_per_day),2)
       median_per_day <<- round(median(steps_per_day),2)
       hist(steps_per_day, col = "cadetblue", main = "Total number of steps taken per day", xlab = "", ylab = "")
       abline(v = mean_per_day, col = "gold", lwd = 3 )
       abline(v = median_per_day, col = "deeppink", lwd = 3 )
       legend("topright", legend = c(paste("mean =",mean_per_day), paste("median =",median_per_day)), 
                                     lwd = 3,col = c("gold","deeppink"))
}

num_steps_per_per_day_hist(dt)
```
### The mean and median of original data:  
### - mean = `r mean_per_day`
### - median = `r median_per_day`
## What is the average daily activity pattern?
```{r daily_activity_pattern, fig.width= 11}
daily_activity_pattern = sapply(split(dt$steps,dt$interval),mean, na.rm = TRUE)
#daily_activity_pattern = as.vector(daily_activity_pattern, mode = "numeric")
max_interval = which.max(daily_activity_pattern)
max_activity = daily_activity_pattern[max_interval]
t = max_interval * 5
h = as.integer(t/60)
m = t -h*60
str_t = paste(h, m, sep = ":")
plot(daily_activity_pattern, type = "l", col = "cadetblue", lwd = 3, main = "Daily activity partern", xlab = "", ylab = "")
abline(v = max_interval, lty = 4)
legend("topright", legend = paste("the most active time is around", str_t), lty = 4)
```
### The most "active" 5-minute interval is `r max_interval`, that corresponds to `r str_t`.  
### What about yours most active time of the day?
  
## Imputing missing values
```{r missing_values, fig.width= 11 }
steps_NA = sum(is.na(dt$steps))
dt_noNA = dt
dt_noNA$steps[is.na(dt$steps)] = daily_activity_pattern[as.character(dt$interval[which(is.na(dt$steps), arr.ind=TRUE)])]
#d = dim(dt)[1]
#for( i in 1 : d ){
#       if (is.na(dt$steps[i]))
#              dt$steps[i] = daily_activity_pattern[as.character(dt$interval[i])]
#}
num_steps_per_per_day_hist(dt_noNA)
```
### There was `r steps_NA` missing values in original data set.
### Missing data was replaced by the mean of corresponding 5-minute interval.
### And as we see, filling missing data makes median and mean equal.


## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays, fig.width= 11}
library(lattice)
dt_noNA$days = weekdays(as.Date(dt_noNA$date,"%Y-%m-%d"))
dt_noNA$weekday = as.factor(dt_noNA$days != "Sunday" & dt_noNA$days != "Saturday")
levels(dt_noNA$weekday) = c("weekend","weekday")
dt_wd = split(dt_noNA,dt_noNA$weekday)
weekend_activity_pattern = sapply(split((dt_wd$weekend)$steps,(dt_wd$weekend)$interval),mean, na.rm = TRUE)
weekday_activity_pattern = sapply(split((dt_wd$weekday)$steps,(dt_wd$weekday)$interval),mean, na.rm = TRUE)
#weekend_activity_pattern = as.vector(weekend_activity_pattern)
#weekday_activity_pattern = as.vector(weekday_activity_pattern)
activity_pattern = append(weekday_activity_pattern,weekend_activity_pattern)
daytime = as.integer(names(activity_pattern))
activity_pattern = as.vector(activity_pattern)
f = c(rep("weekday", length(weekday_activity_pattern)),rep("weekend", length(weekend_activity_pattern)))
f = factor(f)
#df = data.frame( activity = activity_pattern,weekday = f)
xyplot(activity_pattern ~ daytime | f, layout = c(1,2), type = "l", col = "cadetblue",
       xlab = "Interval", ylab = "Number of steps", lwd = 3 )
```
### The patterns look similar, but on weekends less early morning activity and "rush hour" pick.
###  So, weekends are more healthy!